---
title: "Brown et al. temperature summary"
output: html_document
date: "2023-12-22"
editor_options: 
  chunk_output_type: console
---

# Set Up  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library("car") #levenes test
library("ggplot2") #plotting
library("plotrix") #plotting
library("reshape2") #data shaping
require("gridExtra") #Arrange Plots for output
require("plyr") #for ddply
require("dplyr")
require("utils")
library("tidyverse")
library("lubridate")
library("scales")
```

Data downloaded from: https://www.bco-dmo.org/dataset/905047  
GitHub: https://github.com/imkristenbrown/Divergent-bleaching-and-recovery-trajectories-in-corals-following-of-heatwaves   
Publication: https://www.pnas.org/doi/10.1073/pnas.2312104120  

# Read in data and summarize mean temperatures for spawning windows for context for our study 

```{r}
data<-read_csv("Mcap2021/Data/Environmental/brown_etal_905047_v1_infield_temp.csv")
```

Group by year and month and summarize data from all Kaneohe Bay sites.

2013-2023

Non bleaching years 
```{r}
data%>%
  filter(!year=="2014")%>%
  filter(!year=="2015")%>%
  filter(!year=="2019")%>%
  filter(month>4 & month<9)%>%
  group_by(month)%>%
  summarise(mean=mean(Temperature, na.rm=TRUE), sd=sd(Temperature, na.rm=TRUE))
```
In non bleaching years, mean temperatures are 26-27C. 

Bleaching years 
```{r}
data%>%
  filter(year %in% c("2014", "2015", "2019"))%>%
  filter(month>4 & month<9)%>%
  group_by(month)%>%
  summarise(mean=mean(Temperature, na.rm=TRUE), sd=sd(Temperature, na.rm=TRUE))
```
In bleaching years, temperatures are 26-28.5. 

View daily mean maximums for non bleaching years. 
```{r}
data%>%
  filter(!year=="2014")%>%
  filter(!year=="2015")%>%
  filter(!year=="2019")%>%
  filter(month>4 & month<9)%>%
  group_by(day, year, month)%>%
  summarise(max=max(Temperature, na.rm=TRUE))%>%
  filter(!is.infinite(max))%>%
  group_by(month)%>%
  summarise(max=max(max, na.rm=TRUE))
  
```
In non bleaching years, daily max temperatures range from 28.8-30.1. 

View daily mean maximums for bleaching years. 
```{r}
data%>%
  filter(year %in% c("2014", "2015", "2019"))%>%
  filter(month>4 & month<9)%>%
  group_by(day, year, month)%>%
  summarise(max=max(Temperature, na.rm=TRUE))%>%
  filter(!is.infinite(max))%>%
  group_by(month)%>%
  summarise(max=max(max, na.rm=TRUE))
  
```
In  bleaching years, daily max temperatures range from 28.4-30.3. 



