---
title: Larval respirometry (photosynthesis & respiration) analysis and plotting
author: "AS Huffmyer"
date: '2021'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiontic Integration 2021 respirometry data for Pacuta larvae.   

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```


# **Data manipulation**  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("Pacu2021/Output/Respiration/PR/oxygen_P_R_calc.csv") #load data
```

Format data columns.   
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$Temperature<-as.factor(PRdata$Temperature)
PRdata$Plate<-as.factor(PRdata$Plate)
PRdata$Run<-as.factor(PRdata$Run)
```

Calculate a P:R ratio using gross photosynthesis and respiration calculated as GP:R.    
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-abs(PRdata$GP.nmol.org.min)/abs(PRdata$R.nmol.org.min) #calculate ratio with absolute values
```

View outliers. Remove outliers for P:R data.   
```{r}
boxplot(PRdata$R.nmol.org.min)
boxplot(PRdata$P.nmol.org.min)
boxplot(PRdata$GP.nmol.org.min)
boxplot(PRdata$ratio)
PRdata<-PRdata%>%filter(ratio < 4) #filter out outlier for PR ratio
boxplot(PRdata$ratio)
```

# **Plotting**  

## Respiration  

Plot data as a box plot.  
```{r}
r_plot<-PRdata %>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = R.nmol.org.min, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Temperature, group=Temperature), alpha=0.5)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot
```

Plot data as a mean dot plot.  
```{r}
r_plot_b<-PRdata %>%
    group_by(Temperature)%>%
    summarise(mean=mean(R.nmol.org.min), sd=sd(R.nmol.org.min), N=length(R.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=6, alpha=0.8, position = position_dodge(0.3)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.3), size=1, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.1, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot_b
```

## Photosynthesis (Net)  

For photosynthesis filter out runs 7 and 8, these runs were outliers that didn't work due to equipment issues.    

Plot data as a box plot.    
```{r}
p_plot<-PRdata %>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = P.nmol.org.min, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Temperature, group=Temperature), alpha=0.5)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.02, 0.33), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot
```

Plot data as a mean dot plot.    
```{r}
p_plot_b<-PRdata %>%
    group_by(Temperature)%>%
    summarise(mean=mean(P.nmol.org.min), sd=sd(P.nmol.org.min), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=6, alpha=0.8, position = position_dodge(0.3)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.3), size=1, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.005, 0.15), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot_b

```

## Photosynthesis (Gross)  

Plot data as a box plot.    
```{r}
gp_plot<-PRdata %>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = GP.nmol.org.min, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Temperature, group=Temperature), alpha=0.5)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.02, 0.33), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot
```

Plot data as a mean dot plot.    
```{r}
gp_plot_b<-PRdata %>%
    group_by(Temperature)%>%
    summarise(mean=mean(GP.nmol.org.min), sd=sd(GP.nmol.org.min), N=length(GP.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Temperature)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=6, alpha=0.8, position = position_dodge(0.3)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.3), size=1, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    scale_y_continuous(limits=c(-0.005, 0.15), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot_b
```

## P:R Ratio  

Plot data as a box plot.     
```{r}
pr_plot<-PRdata %>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = ratio, group=Temperature)) +
    geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Temperature, group=Temperature), alpha=0.5)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("P:R")))) +
    scale_y_continuous(limits=c(0, 4), labels = scales::number_format(accuracy = 0.1, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot
```

Plot data as a mean dot plot.     
```{r}
pr_plot_b<-PRdata %>%
    group_by(Temperature)%>%
    summarise(mean=mean(ratio), sd=sd(ratio), N=length(ratio), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Temperature)) +
    geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(fill=Temperature, group=Temperature), pch = 21, size=6, alpha=0.8, position = position_dodge(0.3)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Temperature), width=0, linetype="solid", position=position_dodge(0.3), size=1, color="black")+ 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","orange", "red"))+
    scale_color_manual(name="Temperature", values=c("blue","orange", "red"))+
    ylab(expression(bold(paste("P:R")))) +
    scale_y_continuous(limits=c(0, 2), labels = scales::number_format(accuracy = 0.1, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot_b
```

## Generate final figure  

Box plot assembled figure.  
```{r}
full_fig<-plot_grid(r_plot, p_plot, gp_plot, pr_plot, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.2), label_y=1, align="h")

ggsave(filename="Pacu2021/Figures/Respiration/PR/respirometry_fig_boxplot.png", plot=full_fig, dpi=500, width=18, height=6, units="in")
```

Mean plot assembled figure.  
```{r}
full_fig_b<-plot_grid(r_plot_b, p_plot_b, gp_plot_b, pr_plot_b, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.2), label_y=1, align="h")

ggsave(filename="Pacu2021/Figures/Respiration/PR/respirometry_fig_mean.png", plot=full_fig_b, dpi=500, width=18, height=6, units="in")
```

## **Analysis**   

### *Respiration*     

Change data columns to factors.  
```{r}
PRdata$Temperature<-as.factor(PRdata$Temperature)
PRdata$Date<-as.factor(PRdata$Date)
PRdata$Run<-as.factor(PRdata$Run)
```

View respiration data.   

```{r}
hist(PRdata$R.nmol.org.min)
```

Build model and examine for respiration across temperatures.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
Rmodel1<-aov(R.nmol.org.min~Temperature, data=PRdata) 
summary(Rmodel1)
```

Check assumptions of model for residual normality and variance. 

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel1))
leveneTest(residuals(Rmodel1)~Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(Rmodel1, ~Temperature)
pairs(emm)
```

### *Net Photosynthesis*  

Build model and examine for Photosynthesis across temperatures.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
Pmodel1<-aov(P.nmol.org.min~Temperature, data=PRdata) 
summary(Pmodel1)
```

Check assumptions of model for residual normality and variance.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel1))
leveneTest(residuals(Pmodel1)~Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(Pmodel1, ~Temperature)
pairs(emm)
```

### *Gross Photosynthesis*  

Build model and examine for gross photosynthesis across temperatures.         
```{r, results=TRUE, warning=FALSE, message=FALSE}
GPmodel1<-aov(GP.nmol.org.min~Temperature, data=PRdata)
summary(GPmodel1)
```

Check assumptions of model for residual normality and variance.      

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel1))
leveneTest(residuals(GPmodel1)~Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(GPmodel1, ~Temperature)
pairs(emm)
```

### *P:R*  
 
Build model and examine for PR ratio across temperatures.         
```{r, results=TRUE, warning=FALSE, message=FALSE}
PRmodel1<-aov(ratio~Temperature, data=PRdata) 
summary(PRmodel1)
```

Check assumptions of model for residual normality and variance.

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel1))
leveneTest(residuals(PRmodel1)~Temperature, data=PRdata)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(PRmodel1, ~Temperature)
pairs(emm)
```

### *Generate summary*  

Generate summary of all respiration data.  

```{r}
summary<-PRdata%>%
  group_by(Temperature)%>%
  filter(R.nmol.org.min<0)%>%
  filter(P.nmol.org.min>0)%>%
  filter(GP.nmol.org.min>0)%>%
  summarise(N=length(R.nmol.org.min),
            Mean_R=mean(R.nmol.org.min), 
            SE_R=sd(R.nmol.org.min)/sqrt(length(R.nmol.org.min)),
            Mean_P=mean(P.nmol.org.min), 
            SE_P=sd(P.nmol.org.min)/sqrt(length(P.nmol.org.min)),
            Mean_GP=mean(GP.nmol.org.min), 
            SE_GP=sd(GP.nmol.org.min)/sqrt(length(GP.nmol.org.min)),
            Mean_PR=mean(ratio), 
            SE_PR=sd(ratio)/sqrt(length(ratio)))

summary

summary%>%
  write_csv(., "Pacu2021/Output/Respiration/PR/mean_respiration.csv")
```