---
title: Larval stages respirometry (photosynthesis & respiration) analysis and plotting
author: "AS Huffmyer"
date: '2021'
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for respirometry data across *Pocillopora acuta* larval stages (I, II, III).   

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```


# **Data visualization**  

Load data generated from LoLinR rates extraction.      
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("Pacu2021/Output/Respiration/Stages/oxygen_P_R_calc.csv") #load data
```

Format data columns and remove blank wells.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$Temperature<-as.factor(PRdata$Temperature)
PRdata$Stage<-as.factor(PRdata$Stage)
PRdata$Treatment<-as.factor(PRdata$Treatment)
PRdata$Plate<-as.factor(PRdata$Plate)
PRdata$Run<-as.factor(PRdata$Run)
```

Calculate a P:R ratio = gross photosynthesis / respiration. A 1:1 ratio indicates that photosynthesis meets respiratory demand.   
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-abs(PRdata$GP.nmol.org.min)/abs(PRdata$R.nmol.org.min) #calculate ratio with absolute values
#remove outliers detected by values of P:R ratio data
```

View boxplots for each metric. For P:R ratio, remove outliers >15.    
```{r}
boxplot(PRdata$R.nmol.org.min)
boxplot(PRdata$P.nmol.org.min)
boxplot(PRdata$GP.nmol.org.min)
boxplot(PRdata$ratio)
PRdata<-PRdata%>%filter(ratio < 15) #filter out outlier for PR ratio
boxplot(PRdata$ratio)
```

# **Plotting**  

In these plots, there are several factors displayed:  
1) "Treatment" = larval rearing treatment, either high or ambient  
2) "Temperature" = temperature at which respirometry was conducted, either 27C or 31C  
3) "Stage" = the larval stage, either 1 (planula), 2 (metamorphosing), or 3 (polyp)


## Respiration  

Respiration data provides information on organism energetic demand. Here, respiration is shown as the absolute value of oxygen consumed per larva per unit time.  

First, plot data as a box plot.  

```{r}
r_plot<-PRdata %>%
    droplevels()%>%
    group_by(Temperature)%>%
    mutate(abs=abs(R.nmol.org.min))%>%
    ggplot(., aes(x = as.factor(Temperature), y = abs, group=interaction(Treatment, Temperature))) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Treatment, group=interaction(Treatment,Temperature)), alpha=0.5)+
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue","red"))+
    scale_color_manual(name="Temperature", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot
```

Next plot as an interaction plot.  

```{r}
r_plot_b<-PRdata %>%
    group_by(Temperature, Treatment, Stage)%>%
    mutate(abs=abs(R.nmol.org.min))%>%
    summarise(mean=mean(abs), sd=sd(R.nmol.org.min), N=length(R.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Treatment)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.3), size=1.3, alpha=0.75,aes(color=Treatment, group=Treatment)) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Treatment), width=0, linetype="solid", position=position_dodge(0.3), size=0.8, color="black")+
     geom_point(aes(fill=Treatment, group=Treatment), pch = 21, size=6, alpha=0.75, position = position_dodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Treatment", values=c("blue","red"))+
    scale_color_manual(name="Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot_b
```


## Photosynthesis (Net)  

Net photosynthesis shows the oxygen production that occurs in excess of oxygen consumption (respiration).  

Plot data as a box plot.  

```{r}
p_plot<-PRdata %>%
    droplevels()%>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = P.nmol.org.min, group=Temperature)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Treatment, group=interaction(Treatment,Temperature)), alpha=0.5)+
    geom_point(aes(fill=Treatment, group=interaction(Treatment,Temperature)), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue", "red"))+
    scale_color_manual(name="Temperature", values=c("blue", "red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.005, 0.03), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot

```

Plot as an interaction plot.  

```{r}
p_plot_b<-PRdata %>%
    group_by(Temperature, Treatment, Stage)%>%
    summarise(mean=mean(P.nmol.org.min), sd=sd(P.nmol.org.min), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Treatment)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.3), size=1.3, alpha=0.75,aes(color=Treatment, group=Treatment)) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Treatment), width=0, linetype="solid", position=position_dodge(0.3), size=0.8, color="black")+
     geom_point(aes(fill=Treatment, group=Treatment), pch = 21, size=6, alpha=0.75, position = position_dodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Treatment", values=c("blue","red"))+
    scale_color_manual(name="Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot_b
```


## Photosynthesis (Gross)  

Gross photosynthesis is calculated by removing oxygen consumption (respiration) therefore showing total oxygen production (photosynthesis).  

First, lot data as a box plot.  


```{r}
gp_plot<-PRdata %>%
    group_by(Temperature)%>%
    filter(GP.nmol.org.min>0)%>%
    ggplot(., aes(x = as.factor(Temperature), y = GP.nmol.org.min, group=Temperature)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Treatment, group=interaction(Treatment,Temperature)), alpha=0.5)+
    geom_point(aes(fill=Treatment, group=interaction(Treatment,Temperature)), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue", "red"))+
    scale_color_manual(name="Temperature", values=c("blue", "red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.005, 0.05), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot

```

Next plot as an interaction plot.  

```{r}
gp_plot_b<-PRdata %>%
    group_by(Temperature, Treatment, Stage)%>%
    summarise(mean=mean(GP.nmol.org.min), sd=sd(GP.nmol.org.min), N=length(GP.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Treatment)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.3), size=1.3, alpha=0.75,aes(color=Treatment, group=Treatment)) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Treatment), width=0, linetype="solid", position=position_dodge(0.3), size=0.8, color="black")+
     geom_point(aes(fill=Treatment, group=Treatment), pch = 21, size=6, alpha=0.75, position = position_dodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Treatment", values=c("blue","red"))+
    scale_color_manual(name="Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot_b
```

## P:R Ratio  

P:R ratio shows the energy balance between photosynthesis and respiration. A ratio of 1 indicates that gross photosynthesis = respiration. Values >1 indicate more oxygen is produced than consumed.  

Plot data as a box plot.  


```{r}
pr_plot<-PRdata %>%
    group_by(Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = ratio, group=Temperature)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(fill=Temperature, group=interaction(Treatment,Temperature)), alpha=0.5)+
    geom_point(aes(fill=Temperature, group=interaction(Treatment,Temperature)), pch = 21, size=4, alpha=0.5, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Temperature", values=c("blue", "red"))+
    scale_color_manual(name="Temperature", values=c("blue", "red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(0.8, 3), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot

```

Plot as an interaction plot.  
```{r}
pr_plot_b<-PRdata %>%
    group_by(Temperature, Treatment, Stage)%>%
    summarise(mean=mean(ratio), sd=sd(ratio), N=length(ratio), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=Treatment)) +
    facet_wrap(~Stage)+
    geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.3), size=1.3, alpha=0.75,aes(color=Treatment, group=Treatment)) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=Treatment), width=0, linetype="solid", position=position_dodge(0.3), size=0.8, color="black")+
     geom_point(aes(fill=Treatment, group=Treatment), pch = 21, size=6, alpha=0.75, position = position_dodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Treatment", values=c("blue","red"))+
    scale_color_manual(name="Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(-0.2, 0.06), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot_b
```

## Generate final figure of all plots together.     

Box plot figures:  
```{r}
full_fig<-plot_grid(r_plot, p_plot, gp_plot, pr_plot, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1), label_y=1, align="h")

ggsave(filename="Pacu2021/Figures/Respiration/Stages/stages_boxplots.png", plot=full_fig, dpi=500, width=18, height=6, units="in")
```

Interaction plot figures:  
```{r}
full_fig_b<-plot_grid(r_plot_b, p_plot_b, gp_plot_b, pr_plot_b, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1), label_y=1, align="h")

ggsave(filename="Pacu2021/Figures/Respiration/Stages/stages_interaction.png", plot=full_fig_b, dpi=500, width=24, height=6, units="in")
```


# **Analyze data**    

Analyze data using an anova analysis with treatment (larval rearing), temperature (temperature of measurement) and stage as main effects. For these analysis, only include Stage 1 and Stage 2, becuase we do not have high treatment larvae for Stage 3.      

## Respiration      
 
Change variables to factors.  
```{r}
PRdata$Temperature<-as.factor(PRdata$Temperature)
PRdata$Treatment<-as.factor(PRdata$Treatment)
PRdata$Stage<-as.factor(PRdata$Stage)
PRdata$Date<-as.factor(PRdata$Date)
PRdata$Run<-as.factor(PRdata$Run)
```

Build model and examine for respiration.      
```{r, results=TRUE, warning=FALSE, message=FALSE}
model_data<-PRdata%>%
  filter(!Stage=="Stage3")

Rmodel1<-aov(R.nmol.org.min~Temperature*Treatment*Stage, data=model_data)
summary(Rmodel1)
```

Temperature x treatment interaction is significant for respiration.  

Check assumptions of model for residual normality and variance. 

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel1))
leveneTest(residuals(Rmodel1)~Temperature*Treatment*Stage, data=model_data)
```
Passes assumptions.    

## Net Photosynthesis    

Build model and examine for net photosynthesis.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
Pmodel1<-aov(P.nmol.org.min~Temperature*Treatment*Stage, data=model_data) 
summary(Pmodel1)
```

Temperature and stage are significant for net photosynthesis.  

Check assumptions of model for residual normality and variance.     

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel1))
leveneTest(residuals(Pmodel1)~Temperature*Treatment*Stage, data=model_data)
```

Passes assumptions.  

Conduct post hoc test to look at pairwise differences for temperature and treatments between stages.      

```{r}
emm<-emmeans(Pmodel1, ~Temperature*Treatment | Stage)
pairs(emm)
```

Next, conduct post hoc test to look at pairwise differences for stages and treatments between temperatures      

```{r}
emm<-emmeans(Pmodel1, ~Stage*Treatment | Temperature)
pairs(emm)
```

## Gross Photosynthesis    

Build model and examine for gross photosynthesis. Remove outliers of ratio >6.          
```{r, results=TRUE, warning=FALSE, message=FALSE}
GPmodel1<-aov(GP.nmol.org.min~Temperature*Treatment*Stage, data=model_data) 
summary(GPmodel1)
```

Stage and Temperature x treatment interaction are significant.  

Check assumptions of model for residual normality and variance.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel1))
leveneTest(residuals(GPmodel1)~Temperature*Treatment*Stage, data=model_data)
```

Passes. 

Conduct post hoc test looking between stages.  

```{r}
emm<-emmeans(GPmodel1, ~Temperature*Treatment | Stage)
pairs(emm)
```

Conduct post hoc test looking between temperatures  

```{r}
emm<-emmeans(GPmodel1, ~Stage*Treatment | Temperature)
pairs(emm)
```

##  P:R ratio   

Build model and examine for PR ratio.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
PRmodel1<-aov(ratio~Temperature*Treatment*Stage, data=model_data)
summary(PRmodel1)
```

No significant effects for P:R ratio.  

Check assumptions of model for residual normality and variance. 

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel1))
leveneTest(residuals(PRmodel1)~Temperature*Treatment*Stage, data=model_data)
```

Meets assumptions.  

Conduct post hoc test within stage.  

```{r}
emm<-emmeans(PRmodel1, ~Temperature*Treatment | Stage)
pairs(emm)
```

Conduct post hoc test within temperature  

```{r}
emm<-emmeans(PRmodel1, ~Stage*Treatment | Temperature)
pairs(emm)
```

## Generate summary table   

Generate summary of all respiration data.  

```{r}
summary<-PRdata%>%
  group_by(Temperature, Treatment, Stage)%>%
  summarise(N=length(R.nmol.org.min),
            Mean_R=mean(R.nmol.org.min), 
            SE_R=sd(R.nmol.org.min)/sqrt(length(R.nmol.org.min)),
            Mean_P=mean(P.nmol.org.min), 
            SE_P=sd(P.nmol.org.min)/sqrt(length(P.nmol.org.min)),
            Mean_GP=mean(GP.nmol.org.min), 
            SE_GP=sd(GP.nmol.org.min)/sqrt(length(GP.nmol.org.min)),
            Mean_PR=mean(ratio), 
            SE_PR=sd(ratio)/sqrt(length(ratio)));summary

summary%>%
  write_csv(., "Pacu2021/Output/Respiration/Stages/mean_respiration.csv")
```