---
title: Carbohydrate analysis
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiotic Integration 2021 carbohydrate data. 

# **Setup**  

Load libraries. 

```{r}
library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(car)
```

# **Load data**  

Read in datafiles

```{r}
data <- read.csv("Mcap2021/Data/Carbohydrates/20220406_Carbohydrates_data.csv")
Meta <- read.csv("Mcap2021/Data/Carbohydrates/20220406_Carb_Meta.csv")
volumes <- read.csv("Mcap2021/Data/Resuspension_volumes.csv")
```

# **Standardize and calculate**  
## **Run with only Initial Standard curve**  

Merging Files and renaming columns

```{r}
data.1 <- merge(Meta, data, by = c("Well", "Run"))
data.1 <- left_join(data.1, volumes, by=c("Tube.ID"))

# Blank correction for each run separately

Blank <- data.1 %>% 
  filter(Sample.Type == "Blank") %>%
  filter(Run=="Initial")%>%
  summarise(blk.avg = mean(X485))

data.1$abs.corr <- data.1$X485 - Blank$blk.avg
```

```{r}
# Standard curve 

Standard <- data.1 %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run=="Initial")

Standard.plot <- ggplot(data = Standard, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot
```

Extract standard curve.  
```{r}
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 
```

## **Calculations (initial curve)**  

Obtain concentration values for samples from standard curve.  

```{r}
Samples <- data.1 %>% #subsetting Samples
  filter(Sample.Type == "Sample")%>%
  filter(!Run=="Initial")

Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration
```

Accounting for dilution factor (1000/100) and normalizing to homogenate volume 

Multiply sample concentration (mg/mL) by total slurry volume (mL) and dilution factor (1000/v of sample, usually 100 mL)

```{r}
Samples$Carb.mg <- (Samples$Concentration * (Samples$Resuspension_volume/1000) * (1000/Samples$Homo_vol))

#Concentration = mg/mL
#Dilution factor 1000/10 for regular protocol doing 900:100 DI:sample
#For us we have 925:75 or 1000/25

#the second term here needs to be the total amount in the host or holobiont tube that we pulled from 
#needs to be VOLUME from this page: https://docs.google.com/spreadsheets/d/1kE8Yamrzdh5VDtieGuJGIl37WnuAOXeknBqTraglg3k/edit#gid=316770111
#can call resuspension volume
#need to get these values from Jill 

#then the third term is our "input" volume (called Homo_vol here)

```

Carbohydrates are now in total mg carbohydrates in each sample.  

Plot by individual sample

```{r}
Carb.Plot.samples <- Samples%>%
  filter(!Treatment=="Astrangia")%>%
  ggplot(aes(x=Tube.ID, y=Carb.mg, fill = Fraction)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
   #scale_fill_manual(values=c("blue", "red"))+
#  stat_summary(fun=mean, aes(group=Fraction, color = Fraction), position = position_dodge(width = 0.5))  + 
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  #  ylim(0,0.5) +
  #geom_text(position=position_jitterdodge(dodge.width=0.4))+
  xlab("Temperature") + ylab(expression("Total Carbohydrate " (mg))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, linewidth=0.75), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  #  theme(axis.text = element_text(size = 30, color = "black"),
  #        axis.title = element_text(size = 36, color = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Carb.Plot.samples 
```

We need to remove H, since it was a very dilute sample.  

Plot only larvae

```{r}
Carb.Plot <- Samples%>%
  filter(!Treatment=="Astrangia")%>%
  filter(!Sample.ID=="H")%>%
  ggplot(aes(x=Treatment, y=Carb.mg, fill = Fraction)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Temperature") + ylab(expression("Total Carbohydrate " (mg))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), 
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Carb.Plot 
```

Test for differences with an anova.  

```{r}
model<-Samples%>%
  filter(!Treatment=="Astrangia")%>%
  filter(!Sample.ID=="H")%>%
  
  aov(Carb.mg~Treatment*Fraction, data=.)

summary(model)
```

## **Run with all runs Standard curve**  

Merging Files and renaming columns

```{r}
data.1.all <- merge(Meta, data, by = c("Well", "Run"))
data.1.all <- left_join(data.1.all, volumes, by=c("Tube.ID"))

# Blank correction for each run separately

Blank.all <- data.1.all %>% 
  filter(Sample.Type == "Blank") %>%
  #filter(Run=="Initial")%>%
  summarise(blk.avg = mean(X485))

data.1.all$abs.corr <- data.1.all$X485 - Blank.all$blk.avg
```

```{r}
# Standard curve 

Standard.all <- data.1.all %>% 
  filter(Sample.Type == "Standard") #%>%
  #filter(Run=="Initial")

Standard.plot.all <- ggplot(data = Standard.all, aes(x=Concentration, y=abs.corr))+
  facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Carbohydrate (mg/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot.all
```

Extract standard curve.  
```{r}
lmstandard.all <- lm (Concentration ~ abs.corr, data = Standard.all)
lmsummary.all <- summary(lmstandard.all) 
```

## **Calculations (using all curves combined)**  

Obtain concentration values for samples from standard curve.  

```{r}
Samples.all <- data.1.all %>% #subsetting Samples
  filter(Sample.Type == "Sample")%>%
  filter(!Run=="Initial")

Samples.all$Concentration <- predict(lmstandard.all, newdata = Samples.all) #using model to get concentration
```

Accounting for dilution factor (1000/10) and normalizing to homogenate volume 

```{r}
Samples.all$Carb.mg <- (Samples.all$Concentration * (Samples.all$Resuspension_volume/1000) * (1000/Samples.all$Homo_vol))

#Concentration = mg/mL
#Dilution factor 1000/10 for regular protocol doing 900:100 DI:sample
#For us we have 925:75 or 1000/25

#the second term here needs to be the total amount in the host or holobiont tube that we pulled from 
#needs to be VOLUME from this page: https://docs.google.com/spreadsheets/d/1kE8Yamrzdh5VDtieGuJGIl37WnuAOXeknBqTraglg3k/edit#gid=316770111
#can call resuspension volume

#then the third term is our "input" volume (called Homo_vol here)

```

Plot only larvae

```{r}
Carb.Plot.all <- Samples.all%>%
  filter(!Sample.ID=="H")%>%
  filter(!Treatment=="Astrangia")%>%
  ggplot(aes(x=Treatment, y=Carb.mg, fill = Fraction)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Temperature") + ylab(expression("Total Carbohydrate " (mg))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), 
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Carb.Plot.all 
```


## **Show plots side by side**  

```{r}
library(cowplot)

larvae.compare<-plot_grid(Carb.Plot, Carb.Plot.all, nrow=1, ncol=2)
larvae.compare

```

We get the same answer with very minor differences using the curves. So we will proceed with using the first curve (Initial Run) because it was more robust.  

# **Host:holobiont carbohydrate ratios**   

Calculate the ratio of host:holobiont carbohydrates 

Calculate the average host and average holobiont signal for each sample.  
```{r}
ratios<-Samples%>%
  filter(!Treatment=="Astrangia")%>%
  filter(!Sample.ID=="H")%>%
  dplyr::select(Sample.ID, Treatment, Fraction, Carb.mg)%>%
  group_by(Sample.ID, Fraction, Treatment)%>%
  dplyr::summarise(mean_carbs=mean(Carb.mg))%>%
  spread(Fraction, mean_carbs)%>%
  mutate(ratio=Host/Holobiont)
  
```

## Statistical tests 

Test for differences with an anova.  

```{r}
model<-aov(ratio~Treatment, data=ratios)
summary(model)
```

There is a slight but significant difference in Host:Holobiont ratios of carbohydrates at high temperature. There are more carbohydrates in the host fraction out of the total carbohydrates at high temperature. 

## Plotting  

Plot Host:Holobiont ratios  

```{r}
Ratio.Plot <- ratios%>%
  ggplot(aes(x=Treatment, y=ratio)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
   scale_fill_manual(values=c("blue", "red"))+
  scale_colour_manual(values=c("blue", "red"))+
  geom_text(x=1.5, y=0.48, label="p=0.046", color="black", size=4)+
  xlab(expression(bold("Temperature"))) + ylab(expression(bold("Host:Holobiont Carb."))) + #Axis titles
  theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black")); Ratio.Plot 
```

# **Normalize to protein values**  

Read in protein values (total protein per sample in ug).  
```{r}
protein<-read.csv("Mcap2021/Output/Protein/protein_output.csv")
```

Merge data frames.  
```{r}
Samples_norm<-Samples%>%
  filter(!Treatment=="Astrangia")%>%
  filter(!Sample.ID=="H")%>%
  dplyr::group_by(Tube.ID, Sample.ID, Treatment)%>%
  dplyr::summarise(Carb.mg=mean(Carb.mg))

Samples_norm<-left_join(Samples_norm, protein)
```

Convert carbs from mg to ug (multiply by 1000). Normalize carbs (ug) to protein (ug) to obtain carbs per unit protein.  
```{r}
Samples_norm<-Samples_norm%>%
  mutate(Carb.ug=Carb.mg*1000)%>%
  mutate(Carb.normalized=Carb.ug/Protein.ug)
```

## Statistical tests  

Test for differences with an anova for host carbohydrates.  

```{r}
model_data<-Samples_norm%>%
  filter(Fraction=="Host")

model<-aov(Carb.normalized~Treatment, data=model_data)
summary(model)
```
There is no significant difference in host carbohydrates. 

Test for differences with an anova for holobiont carbohydrates.  

```{r}
model_data<-Samples_norm%>%
  filter(Fraction=="Holobiont")

model<-aov(Carb.normalized~Treatment, data=model_data)
summary(model)
```
There is no significant difference in host carbohydrates. 

## Plotting  

Plot host and holobiont fractions. 

```{r}
Carb.Plot.norm <- Samples_norm%>%
  ggplot(aes(x=Treatment, y=Carb.normalized, fill = Fraction)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Temperature") + ylab(expression("Total Carb. " (ug/ug))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), 
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Carb.Plot.norm 
```

Plot host carbohydrates across temperatures.   

```{r}
carb.host.prot <- Samples_norm%>%
  filter(Fraction=="Host")%>%
  
  ggplot(aes(x=Treatment, y=Carb.normalized)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
   scale_fill_manual(values=c("blue", "red"))+
  scale_colour_manual(values=c("blue", "red"))+
  geom_text(x=1.5, y=8, label="p=0.078", color="darkgray", size=4)+
  xlab(expression(bold("Temperature"))) + ylab(expression(bold(paste("Host Carb. (ug ug protein"^-1, ")"))))+ #Axis titles
  theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black")); carb.host.prot 
```

Plot holobiont carbohydrates across temperatures.   

```{r}
carb.holobiont.prot <- Samples_norm%>%
  filter(Fraction=="Holobiont")%>%
  
  ggplot(aes(x=Treatment, y=Carb.normalized)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
   scale_fill_manual(values=c("blue", "red"))+
  scale_colour_manual(values=c("blue", "red"))+
  geom_text(x=1.5, y=15, label="p=0.258", color="darkgray", size=4)+
  xlab(expression(bold("Temperature"))) + ylab(expression(bold(paste("Holobiont Carb. (ug ug protein"^-1, ")"))))+ #Axis titles
  theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black")); carb.holobiont.prot 
```


# **Normalize to cell density**  

## Holobiont    

Read in cell density values.  
```{r}
cells<-read.csv("Mcap2021/Output/Cells/cells_output.csv")
```

Merge data frames.  
```{r}
cells_data_holo<-Samples%>%
  dplyr::select(Treatment, Fraction, Sample.ID, Resuspension_volume, Carb.mg)%>%
  filter(!Treatment=="Astrangia")%>%
  filter(Fraction=="Holobiont")%>%
  filter(!Sample.ID=="H")%>%
  group_by(Treatment, Fraction, Sample.ID, Resuspension_volume)%>%
  dplyr::summarise(Carb.mg=mean(Carb.mg))%>%
  mutate(Carb.mg.uL=Carb.mg/Resuspension_volume)

cells<-cells%>%
  dplyr::select(Sample.ID, cells.uL)

cells_data_holo<-left_join(cells_data_holo, cells, by="Sample.ID")

cells_data_holo<-cells_data_holo%>%
  mutate(Carb.mg.cell=Carb.mg.uL/cells.uL)%>%
  mutate(Carb.ug.cell=Carb.mg.cell*1000)
```

### Statistical tests  

Test for differences with an anova.  

```{r}
cells_data_model_holo<-cells_data_holo%>%
  filter(!Sample.ID=="H")

model<-aov(Carb.ug.cell~Treatment, data=cells_data_model_holo)
summary(model)
```

There are less carbohydrates in the holobiont per symbiont cell. 

### Plotting 

Plot  

```{r}
Carb.Plot.cells.holo <- cells_data_holo%>%
  filter(!Sample.ID=="H")%>%
  
  ggplot(aes(x=Treatment, y=Carb.ug.cell)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
   scale_fill_manual(values=c("blue", "red"))+
  scale_colour_manual(values=c("blue", "red"))+
  geom_text(x=1.5, y=0.034, label="p=0.048", color="black", size=4)+
  xlab(expression(bold("Temperature"))) + ylab(expression(bold(paste("Holobiont Carb. (ug cell"^-1, ")")))) + #Axis titles
  theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black")); Carb.Plot.cells.holo 
```


## Host  

Read in cell density values.  
```{r}
cells<-read.csv("Mcap2021/Output/Cells/cells_output.csv")
```

Merge data frames.  
```{r}
cells_data_host<-Samples%>%
  dplyr::select(Treatment, Fraction, Sample.ID, Resuspension_volume, Carb.mg)%>%
  filter(!Treatment=="Astrangia")%>%
  filter(Fraction=="Host")%>%
  filter(!Sample.ID=="H")%>%
  group_by(Treatment, Fraction, Sample.ID, Resuspension_volume)%>%
  dplyr::summarise(Carb.mg=mean(Carb.mg))%>%
  mutate(Carb.mg.uL=Carb.mg/Resuspension_volume)

cells<-cells%>%
  dplyr::select(Sample.ID, cells.uL)

cells_data_host<-left_join(cells_data_host, cells, by="Sample.ID")

cells_data_host<-cells_data_host%>%
  mutate(Carb.mg.cell=Carb.mg.uL/cells.uL)%>%
  mutate(Carb.ug.cell=Carb.mg.cell*1000)
```

### Statistical Tests  

Test for differences with an anova.  

```{r}
cells_data_model_host<-cells_data_host%>%
  filter(!Sample.ID=="H")
  #filter(Carb.ug.cell<0.02)%>%
  #filter(Carb.ug.cell>0.003)

model<-aov(Carb.ug.cell~Treatment, data=cells_data_model_host)
summary(model)
```

No significant difference in host carbohydrates per symbiont cell. 

### Plotting  

Plot   

```{r}
Carb.Plot.cells.host <- cells_data_host%>%
  filter(!Sample.ID=="H")%>%
  
  ggplot(aes(x=Treatment, y=Carb.ug.cell)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(aes(fill=Treatment), pch = 21, position=position_jitterdodge(dodge.width=0.4), size=4, alpha=0.5) +
   scale_fill_manual(values=c("blue", "red"))+
  scale_colour_manual(values=c("blue", "red"))+
  geom_text(x=1.5, y=0.012, label="p=0.769", color="darkgray", size=4)+
  xlab(expression(bold("Temperature"))) + ylab(expression(bold(paste("Host Carb. (ug cell"^-1, ")")))) + #Axis titles
  theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black")); Carb.Plot.cells.host 
```


# **Assemble figure panel**  

Ratio.Plot
carb.holobiont.prot
carb.host.prot
Carb.Plot.cells.holo
Carb.Plot.cells.host

Generate a figure with panel of carbohydrate metrics.  
```{r}
Ratio.Plot_panel<-Ratio.Plot+theme(axis.title.x=element_blank(), legend.title=element_text(face="bold", size=14), legend.text=element_text(size=14, color="black"))
carb.holobiont.prot_panel<-carb.holobiont.prot+theme(legend.position="none", axis.title.x=element_blank())
carb.host.prot_panel<-carb.host.prot+theme(legend.position="none", axis.title.x=element_blank())
Carb.Plot.cells.holo_panel<-Carb.Plot.cells.holo+theme(legend.position="none", axis.title.x=element_blank())
Carb.Plot.cells.host_panel<-Carb.Plot.cells.host+theme(legend.position="none", axis.title.x=element_blank())

carbs.plot<-plot_grid(carb.holobiont.prot_panel, carb.host.prot_panel, Carb.Plot.cells.holo_panel, Carb.Plot.cells.host_panel, Ratio.Plot_panel, nrow=1, ncol=5, label_y=1, labels=c("A", "B", "C", "D", "E"), label_size=18, rel_widths = c(1,1,1,1,1.4))

ggsave(carbs.plot, file="Mcap2021/Figures/Carbs_panel.png", width=15, height=5)
```

