---
title: "Cell Count Analysis"
author: "AS Huffmyer"
output:
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Script to analyze symbiont density for the Mcap 2021 Symbiont Integration project.  

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("car")) install.packages("car")
if (!require("emmeans")) install.packages("emmeans")
if (!require("multcomp")) install.packages("multcomp")
if (!require("knitr")) install.packages("knitr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("lme4")) install.packages("lme4")
if (!require("GGally")) install.packages("GGally")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(tidyverse)
library(dplyr)
```

# **Symbiont Density**    

Import and manipulate data  
```{r}
# Cell count data
sym_counts <- read.csv("Mcap2021/Data/Cells/cell_counts.csv")

```

Calculate cells and normalize to total protein.  
```{r}
# Calculate mean counts for each sample
df <- sym_counts %>%
  dplyr::select(TubeID, Squares_counted, Sample.ID, matches("Count[1-6]")) %>%
  gather("rep", "count", -TubeID, -Squares_counted, -Sample.ID) %>%
  group_by(TubeID, Squares_counted, Sample.ID) %>%
  dplyr::summarise(mean_count = mean(count, na.rm = TRUE))


#Load in total volume of sample  
volumes<-read.csv("Mcap2021/Data/Resuspension_volumes.csv")
volumes<-dplyr::rename(volumes, TubeID=Tube.ID)

df<-left_join(df, volumes)

# Normalize counts by homogenate volume (ul)
df <- df %>%
  mutate(cells.mL = (mean_count * 10000) / Squares_counted,
         cells = cells.mL * (Resuspension_volume/1000))
        
```

This generated the total number of cells in each tube. 

Now normalized to total protein.   

First, calculate protein per uL from the "Holo" tube for each sample. 

```{r}
protein<-read.csv("Mcap2021/Output/Protein/protein_output.csv")

protein<-protein%>%
  dplyr::rename(TubeID=Tube.ID)

protein<-left_join(protein, volumes)

protein$protein.ug.uL<-protein$Protein.ug/protein$Resuspension_volume

protein<-protein%>%
  filter(Fraction=="Holobiont")
```

Second, divide the total number of cells by the resuspension volume to obtain cells per uL 

```{r}
df$cells.uL=df$cells/df$Resuspension_volume
```

Third, normalize cells per uL / protein per uL to obtain cells per ug protein

```{r}
all<-left_join(df, protein, by="Sample.ID")

all$cells.ugprotein=all$cells.uL/all$protein.ug.uL
```

Output data frame.  

```{r}
all%>%
  dplyr::select(TubeID.x, Sample.ID, cells, cells.uL, cells.ugprotein, Resuspension_volume.x, Treatment)%>%
  dplyr::rename(Tube.ID=TubeID.x, Resuspension_volume=Resuspension_volume.x)%>%
  write_csv("Mcap2021/Output/Cells/cells_output.csv")
```

### Plotting  

Plot cells per unit protein data with mean and standard error.   

```{r}
all %>%
  filter(!Sample.ID=="H")%>%
  ggplot(aes(x = Treatment, y = cells.ugprotein, color = Treatment)) +
  labs(x = "",y = "Cell Density per ug protein") +
  scale_color_manual(values=c("blue", "red"))+
  geom_jitter(width = 0.05) +                                            # Plot all points
  geom_text(aes(label=Sample.ID), vjust = -0.5, hjust=-0.1)+
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun = mean, geom = "point", color = "black") + # Plot mean
  theme_classic()
```


Present means and standard error of each group and save summary table.  

```{r}
all%>%
  group_by(Treatment)%>%
    filter(!Sample.ID=="H")%>%
  summarise(n=length(cells.ugprotein),
            Mean=format(round(mean(cells.ugprotein), 0), 0), 
            SE=format(round(sd(cells.ugprotein)/sqrt(length(cells.ugprotein)),0),0))%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 
```

Plot data as a scatterplot with labels for each sample.    
```{r}
symb_plot<-all %>%
    filter(!Sample.ID=="H")%>%
    droplevels()%>%
    ggplot(., aes(x = as.factor(Treatment), y = cells.ugprotein)) +
    geom_point(aes(fill=Treatment, group=Treatment), pch = 21, size=4, position = position_jitterdodge(0.1)) + 
    geom_text(aes(label=Sample.ID), vjust = -0.5, hjust=-0.1)+
    xlab("Treatment") + 
    #scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    ylab(expression(bold(paste("Symbiont cells ug protein"))))+
    #ylim(0,9000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); symb_plot

```

Plot data as box plot  
```{r}
symb_plot2<-all %>%
    filter(!Sample.ID=="H")%>%
    ggplot(., aes(x = as.factor(Treatment), y = cells.ugprotein)) +
    geom_boxplot(aes(color=Treatment), outlier.size = 0, lwd=1) +
    #geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95, color="black") +
    geom_point(aes(fill=Treatment), pch = 21, size=4, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values=c("blue", "red"))+
  scale_color_manual(values=c("blue", "red"))+
    geom_text(aes(label=Sample.ID), vjust = -0.5, hjust=-0.1)+
    xlab("Treatment") + 
    #scale_fill_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"))+
    #scale_color_manual(name="Lifestage", values=c("#8C510A", "#DFC27D","#80CDC1", "#003C30"), guide="none")+
    ylab(expression(bold(paste("Symbiont cells ug protein"))))+
    #ylim(0,9000)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); symb_plot2

```

### Statistical analysis  

Run ANOVA on cells per protein treatment.    
```{r}
sym_model_data<-all%>%
      filter(!Sample.ID=="H")%>%
      droplevels()

sym_model<-aov(cells.ugprotein~Treatment, data=sym_model_data)
summary(sym_model)
qqPlot(residuals(sym_model))
leveneTest(residuals(sym_model)~Treatment, data=sym_model_data)
```

Both normality and homogeneity of variance pass.   

No significant difference.  


