---
title: "Analyzing larval daily measurement data"
author: "Ariana S Huffmyer"
date: "2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from daily measurements. We are using this for pH and salinity. Temperature is analyzed from temperature loggers in the Temperature_analysis.Rmd script.  

Data was taken multiple times throughout one treatment period day.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Daily measurements  

## Read in file  

Read in daily measurements file.  

```{r}
daily<-read_csv("Mcap2021/Data/Environmental/Daily_Measurements/daily_measurements.csv")
daily$Tank<-as.factor(daily$Tank)
daily$Treatment<-as.factor(daily$Treatment)
daily$tris.date<-as.character(daily$tris.date)
```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("Mcap2021/Data/Environmental/Daily_Measurements/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = Temp.C*Ttris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = pH(Ex=pH.mV, Etris=mVTris, S=Sal.ppt, T=Temp.C)) # calculate pH of the tanks using the pH seacarb function
```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  
```{r}
daily_calc<-pHSlope%>%
  select(Date, Tank, Treatment, Time, Temp.C, Sal.ppt, pH.total, Light.PAR)
```

## Change to long format

Change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=Temp.C:Light.PAR,
  names_to = "metric",
  values_to = "value")
```

## Plot metrics of interest  

Plot by tank colored by tank.      
```{r}
daily_tanks<-daily_calc.long %>%
  ggplot(aes(x=Tank, y=value, colour=Treatment))+
  geom_point(size=2, position=position_dodge(0.3))+
  xlab("Tank")+
  facet_grid(metric ~ ., scales = "free")+
  theme_classic(); daily_tanks

```

## Summarize daily measurements  

Calculate descriptive statistics   
```{r}
daily_calc.short<-daily_calc.long%>%
  pivot_wider(names_from=metric, values_from=value)

summary<-daily_calc.short%>%
  select(!Date)%>%
  select(!Time)%>%
  select(!Tank)%>%
  group_by(Treatment)%>%
  dplyr::summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

summary<-daily_calc.short%>%
  select(!Date)%>%
  select(!Time)%>%
  select(!Tank)%>%
  select(!Treatment)%>%
  dplyr::summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

length(daily_calc.short$Temp.C)
```
