---
title: Size analysis 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiotic Integration 2021 larval size 

# Setup  

Load libraries. 

```{r}
library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(lme4)
library(lmerTest)
library(car)
library(emmeans)
```

# Load data  

```{r}
size<-read_csv("Mcap2021/Data/Size/larval_size.csv")

size<-size%>%
  select(!volume.mm3)
```

# Calculate size 

Calculate volume as a prolate spheroid. 
```{r}
size<-size%>%
  mutate(volume.mm3=((4/3) * pi * (length.mm/2) * (width.mm/2)^2))
```

# Visualize  

Visualize by timepoint and treatment.  

```{r}
size_plot<-size %>%
    ggplot(., aes(x = as.factor(timepoint), y = volume.mm3)) +
    geom_boxplot(aes(color=treatment), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=treatment), pch = 21, size=4, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values=c("blue", "red"))+
    scale_color_manual(values=c("blue", "red"))+
    xlab("Timepoint") + 
    ylab(expression(bold(paste("Larval Size (mm"^3, ")"))))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot
```

Summarize by tube.  

```{r}
size_plot_2<-size %>%
    group_by(tube.ID, treatment, tank, timepoint)%>%
    dplyr::summarise(mean=mean(volume.mm3))%>%
  
    ggplot(., aes(x = as.factor(timepoint), y = mean)) +
    geom_boxplot(aes(color=treatment), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=treatment), pch = 21, size=4, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values=c("blue", "red"))+
    scale_color_manual(values=c("blue", "red"))+
    xlab("Timepoint") + 
    ylab(expression(bold(paste("Larval Size (mm"^3, ")"))))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot_2
```

Larvae grow over the course of the experiment.  

Visualize just at the end.  

```{r}
size_plot_3a<-size %>%
    filter(timepoint=="Sampling")%>%
  
    ggplot(., aes(x = as.factor(treatment), y = volume.mm3)) +
    geom_boxplot(aes(color=treatment), outlier.shape=NA, lwd=0.5) +
    geom_point(aes(fill=treatment), pch = 21, size=4, position = position_jitterdodge(0.2), alpha=0.2) + 
    scale_fill_manual(values=c("blue", "red"))+
    scale_color_manual(values=c("blue", "red"))+
    xlab("") + 
    ylab(expression(bold(paste("Larval Size (mm"^3, ")"))))+
    geom_text(x=1.5, y=0.0, label="p=0.041", color="black", size=5)+
    theme_classic() + 
    ylim(0,0.09)+
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot_3a


#Save plot
ggsave(size_plot_3a, file="Mcap2021/Figures/size_final.png", w=3, h=5)
```

Plot with points and std. error bars.  
```{r}
size_plot_3b<-size %>%
    filter(timepoint=="Sampling")%>%
    group_by(treatment)%>%
    dplyr::summarise(mean=mean(volume.mm3), se=sd(volume.mm3)/sqrt(length(volume.mm3)))%>%

    ggplot(., aes(x = as.factor(treatment), y = mean)) +
    geom_point(aes(fill=treatment), pch = 21, size=8) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=treatment), pch = 21, size=2, width=0) + 
    scale_fill_manual(name="Treatment", values=c("blue", "red"))+
    scale_color_manual(name="Treatment", values=c("blue", "red"))+
    xlab("Treatment") + 
    ylim(0.0, 0.04)+
    ylab(expression(bold(paste("Larval Size (mm"^3, ")"))))+
    geom_text(x=1.5, y=0.01, label="p=0.041", color="black", size=8)+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot_3b
```

Group by tube at the end time point.  
```{r}
size_plot_4<-size %>%
    group_by(tube.ID, treatment, tank, timepoint)%>%
    dplyr::summarise(mean=mean(volume.mm3))%>%
    filter(timepoint=="Sampling")%>%
  
    ggplot(., aes(x = as.factor(treatment), y = mean)) +
    geom_boxplot(aes(color=treatment), outlier.size = 0, lwd=1) +
    geom_point(aes(fill=treatment), pch = 21, size=4, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values=c("blue", "red"))+
    scale_color_manual(values=c("blue", "red"))+
    xlab("Treatment") + 
    ylab(expression(bold(paste("Larval Volume (mm"^3, ")"))))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14)
      ); size_plot_4

```

# Linear models  

Run anova model between time points.  
```{r}
model1<-aov(volume.mm3~timepoint*treatment, data=size)

summary(model1)
qqPlot(residuals(model1))
```

Size is different between treatments.  

Run linear mixed effect model at the ending time points.  
```{r}
model_data<-size%>%
  filter(timepoint=="Sampling")

model2<-aov(volume.mm3~treatment, data=model_data)
summary(model2)

t.test(model_data$volume.mm3~model_data$treatment)

qqPlot(residuals(model2))
```

 
# Final plotting 

Generate custom theme. 
```{r}
custom_theme<-theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        legend.text=element_text(size=12, color="black"), 
        legend.title=element_text(size=12, face="bold", color="black"))
```

Make baseline a gray color instead of treatment color. 
```{r}
size<-size%>%
  mutate(treatment=if_else(timepoint=="Baseline", "Baseline", treatment))
```

Plot. 
```{r}
#make segment locations 
segpoints<-size%>%
  group_by(treatment)%>%
  dplyr::summarise(mean=mean(volume.mm3, na.rm=TRUE))%>%
  spread(treatment, mean)
  
size_plot_full<-size %>%
    ggplot(., aes(x = as.factor(timepoint), y = volume.mm3)) +
    geom_point(aes(fill=treatment), pch = 21, size=4, position = position_jitterdodge(0.6), alpha=0.1) + 
    #geom_line()+
    scale_fill_manual(values=c("blue", "gray", "red"), name="Treatment")+
    scale_color_manual(values=c("blue", "gray", "red"), name="Treatment")+
    ylim(0,0.075)+
    xlab("Timepoint") + 
    ylab(expression(bold(paste("Larval Size (mm"^3, ")"))))+
    geom_text(x=1.5, y=0.069, label="time p=0.070", color="darkgray", size=4)+
    geom_text(x=1.5, y=0.066, label="treatment p=0.047", color="black", size=4)+
  
  #baseline to ambient
    geom_segment(aes(x = 1, y = Baseline, xend = 2, yend = Ambient), colour="blue", data = segpoints, size=2, show.legend=FALSE) +
 
    #baseline to high
    geom_segment(aes(x = 1, y = Baseline, xend = 2, yend = High), colour="red", data = segpoints, size=2, show.legend=FALSE) +
    
    custom_theme+
  theme(axis.title.x=element_blank(), 
        legend.position="none"); size_plot_full

#Save plot
ggsave(size_plot_full, file="Mcap2021/Figures/size_time.png", w=3, h=5)
```

Generate data frame with mean size values for each treatment to use for normalizing metabolic rates.  

```{r}
size %>%
    filter(!treatment=="Baseline")%>%
    group_by(treatment)%>%
    dplyr::summarise(mean.vol.mm3=mean(volume.mm3))%>%
    write_csv("Mcap2021/Output/mean_size.csv")
```

