---
title: Protein analysis
author: "AS Huffmyer"
date: '2022'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiotic Integration 2021 survival data. 

# **Setup**  

Load libraries. 

```{r}
library(dbplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
```

# **Load data**  

Read in datafiles

```{r}
data <- read.csv("Mcap2021/Data/Protein/20220407_Protein_data.csv")
Meta <- read.csv("Mcap2021/Data/Protein/20220407_Protein_Meta.csv")
volumes <- read.csv("Mcap2021/Data/Resuspension_volumes.csv")
```

# **Prepare dataframes**  

Merging Files and renaming columns

```{r}
data.1 <- merge(Meta, data, by = c("Well", "Run"))
data.1 <- left_join(data.1, volumes, by=c("Tube.ID"))

# Blank correction for each run separately

Blank <- data.1 %>% 
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X562))

data.1$abs.corr <- data.1$X562 - Blank$blk.avg
```

Standard for Run 1
```{r}
# Standard curve 

Standard.1 <- data.1 %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==1)

Standard.plot.1 <- ggplot(data = Standard.1, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot.1
```

Extract standard curve for Run 1.  
```{r}
lmstandard.1 <- lm (Concentration ~ abs.corr, data = Standard.1)
lmsummary.1 <- summary(lmstandard.1) 
```

Standard for Run 2
```{r}
# Standard curve 
Standard.2 <- data.1 %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==2)

Standard.plot.2 <- ggplot(data = Standard.2, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot.2
```

Extract standard curve for Run 2.  
```{r}
lmstandard.2 <- lm (Concentration ~ abs.corr, data = Standard.2)
lmsummary.2 <- summary(lmstandard.2) 
```

Standard for Run 3
```{r}
# Standard curve 
Standard.3 <- data.1 %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==3)

Standard.plot.3 <- ggplot(data = Standard.3, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot.3
```

Extract standard curve for Run 3.  
```{r}
lmstandard.3 <- lm (Concentration ~ abs.corr, data = Standard.3)
lmsummary.3 <- summary(lmstandard.3) 
```

There is slight variation in our standard curves. Generate a curve with all run data combined. Later make a script to pull out   

```{r}
# Standard curve 
Standard <- data.1 %>% 
  filter(Sample.Type == "Standard")

Standard.plot <- ggplot(data = Standard, aes(x=Concentration, y=abs.corr))+
  #facet_grid(~Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));Standard.plot
```

Extract standard curve for Run 3.  
```{r}
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 
```

For now, use the combined standard and in later code we will run with individual runs.  

# **Calculations**  

Obtain concentration values for samples from standard curve.  

```{r}
Samples <- data.1 %>% #subsetting Samples
  filter(Sample.Type == "Sample")

Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration
```


Normalizing to homogenization volume WORKING ON THIS

Need to account for dilution (here it will be 25/10) - normally 25 but we used 10
Also need to account for resuspension volume (see carbs script)

Concentration (ug/mL) * resuspension (mL) * dilution factor (ratio)= total ug protein -> this will give us ug per tube, we will then normalize ug carbs to ug protein 

NOte here that standard is in ug/mL

```{r}
Samples$Protein.ug <- Samples$Concentration * (Samples$Resuspension_volume / 1000) * (25/Samples$Homo_vol)

#ug/mL concentration * (resuspension uL / 1000) * (25/homogenate volume added)

```

Plot only larvae

```{r}
Protein.Plot <- Samples%>%
  filter(!Treatment=="Astrangia")%>%
  ggplot(aes(x=Treatment, y=Protein.ug, fill = Fraction)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
#  stat_summary(fun=mean, aes(group=Fraction, color = Fraction), position = position_dodge(width = 0.5))  + 
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  #  ylim(0,0.5) +
  #geom_text(position=position_jitterdodge(dodge.width=0.4))+
  xlab("Temperature") + ylab(expression("Total Carbohydrate " (ug))) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  #  theme(axis.text = element_text(size = 30, color = "black"),
  #        axis.title = element_text(size = 36, color = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Protein.Plot 
```

Test for differences with an anova.  

```{r}
model<-Samples%>%
  filter(!Treatment=="Astrangia")%>%
  
  aov(Protein.ug~Treatment, data=.)

summary(model)
```

# **Calculate Ratio of host:holobiont**   

Calculate the ratio of host:holobiont protein 

Calculate the average host and average holobiont signal for each sample.  
```{r}
ratios<-Samples%>%
  filter(!Treatment=="Astrangia")%>%
  select(Sample.ID, Treatment, Fraction, Protein.ug)%>%
  group_by(Sample.ID, Fraction, Treatment)%>%
  dplyr::summarise(mean_protein=mean(Protein.ug))%>%
  spread(Fraction, mean_protein)%>%
  mutate(ratio=Host/Holobiont)
  
```


Plot Host:Holobiont ratios  

```{r}
Ratio.Plot <- ratios%>%
  ggplot(aes(x=Treatment, y=ratio, fill=Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
#  stat_summary(fun=mean, aes(group=Fraction, color = Fraction), position = position_dodge(width = 0.5))  + 
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  scale_fill_manual(values=c("lightblue", "red1"))+
  #  ylim(0,0.5) +
  #geom_text(position=position_jitterdodge(dodge.width=0.4))+
  xlab("Temperature") + ylab(expression("Host:Holobiont Protein")) + #Axis titles
  theme_bw() + theme(panel.border = element_rect(color="black", fill=NA, size=0.75), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_blank()) +
  #  theme(axis.text = element_text(size = 30, color = "black"),
  #        axis.title = element_text(size = 36, color = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); Ratio.Plot 
```

Test for differences with an anova.  

```{r}
model<-aov(ratio~Treatment, data=ratios)
summary(model)
```

# **Output file to use for protein normalization**  

```{r}
Samples%>%
  filter(!Treatment=="Astrangia")%>%
  select(Tube.ID, Sample.ID, Treatment, Fraction, Protein.ug)%>%
  group_by(Tube.ID, Sample.ID, Treatment, Fraction)%>%
  summarise(Protein.ug=mean(Protein.ug))%>%
  write.csv(., "Mcap2021/Output/Protein/protein_output.csv")
```

