---
title: Respirometry Plotting and Analysis  
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiotic Integration 2021 respirometry data. 

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```

# **Data visualization and manipulation**  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
PRdata<-read.csv("Mcap2021/Output/Respiration/oxygen_P_R_calc.csv") #load data
```

Separate project specific data.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
PRdata<-PRdata[!is.na(PRdata$Type),]

#format columns
PRdata$Temperature<-as.factor(PRdata$Temperature)
PRdata$Plate<-as.factor(PRdata$Plate)
PRdata$Run<-as.factor(PRdata$Run)
```

```{r}
hist(PRdata$R.nmol.org.min)

hist(PRdata$GP.nmol.org.min)
```

If gross photosynthesis is <0, substitute with 0. If respiration is >0, substitute with 0. 
```{r}
PRdata<-PRdata%>%
  mutate(R.nmol.org.min=if_else(R.nmol.org.min>=0, 0, R.nmol.org.min))%>%
  mutate(GP.nmol.org.min=if_else(GP.nmol.org.min<=0, 0, GP.nmol.org.min))%>%
  mutate(P.nmol.org.min=if_else(P.nmol.org.min<=0, 0, P.nmol.org.min))
```

Calculate a P:R ratio using gross photosynthesis and inverse values of respiration.  
```{r, warning=FALSE, message=FALSE}
PRdata$ratio<-PRdata$GP.nmol.org.min/(PRdata$R.nmol.org.min*-1) #calculate ratio with positive values
boxplot(PRdata$ratio)
```

We will need to remove the outlier at 0 and the outliers above 4.
```{r}
PRdata<-PRdata%>%
  mutate(ratio=if_else(ratio>4, NA, ratio))%>%
  mutate(ratio=if_else(ratio<0.1, NA, ratio))
```


# **Analyze data**    

## Individual normalized data

### Respiration     

Run ANOVA.  
```{r}
Rmodel2<-PRdata%>%
  filter(!is.na(R.nmol.org.min))%>%
  
  aov(R.nmol.org.min~Treatment*Temperature, data=.)

summary(Rmodel2)
```

Check assumptions of model for residual normality and variance. Passes assumptions.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel2))

PRdata%>%
  filter(!is.na(R.nmol.org.min))%>%

  leveneTest(residuals(Rmodel2)~Treatment*Temperature, data=.)
```

### Net Photosynthesis   

Run ANOVA.  
```{r}
Pmodel2<-PRdata%>%
  filter(!is.na(P.nmol.org.min))%>%
  
  aov(P.nmol.org.min~Treatment*Temperature, data=.)

summary(Pmodel2)
```

Check assumptions of model for residual normality and variance.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel2))

PRdata%>%
  filter(!is.na(P.nmol.org.min))%>%
  
  leveneTest(residuals(Pmodel2)~Treatment*Temperature, data=.)
```

### Gross Photosynthesis  

Run ANOVA.  
```{r}
GPmodel2<-PRdata%>%
  filter(!is.na(GP.nmol.org.min))%>%
  
  aov(GP.nmol.org.min~Treatment*Temperature, data=.)

summary(GPmodel2)
```

Check assumptions of model for residual normality and variance. Passes.    
```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel2))

PRdata%>%
  filter(!is.na(GP.nmol.org.min))%>%
  
  leveneTest(residuals(GPmodel2)~Treatment*Temperature, data=.)
```

### P:R    
 
```{r, results=TRUE, warning=FALSE, message=FALSE}
hist(PRdata$ratio)
```

Run ANOVA.  
```{r}
PRmodel2<-PRdata%>%
  filter(!is.na(ratio))%>%
  
  aov(ratio~Treatment*Temperature, data=.)

summary(PRmodel2)
```

Check assumptions of model for residual normality and variance. Meets assumptions.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel2))

PRdata%>%
  filter(!is.na(ratio))%>%
  
  leveneTest(residuals(PRmodel2)~Treatment*Temperature, data=.)
```

### Generate summary tables of mean and SE for all variables  

Generate summary of all respiration data.  

```{r}
summary<-PRdata%>%
  group_by(Temperature, Treatment)%>%
  summarise(N=length(R.nmol.org.min),
            Mean_R=mean(R.nmol.org.min*-1, na.rm=TRUE), 
            SE_R=sd(R.nmol.org.min, na.rm=TRUE)/sqrt(length(R.nmol.org.min)),
            Mean_P=mean(P.nmol.org.min, na.rm=TRUE), 
            SE_P=sd(P.nmol.org.min, na.rm=TRUE)/sqrt(length(P.nmol.org.min)),
            Mean_GP=mean(GP.nmol.org.min, na.rm=TRUE), 
            SE_GP=sd(GP.nmol.org.min, na.rm=TRUE)/sqrt(length(GP.nmol.org.min)),
            Mean_PR=mean(ratio, na.rm=TRUE), 
            SE_PR=sd(ratio, na.rm=TRUE)/sqrt(length(ratio)))

summary%>%
  write_csv(., "Mcap2021/Output/Respiration/mean_respiration.csv")
```

## Size normalized data

Read in size data.  

```{r}
size<-read_csv("Mcap2021/Output/mean_size.csv")
```

Calculate a new column for size normalized data by dividing metabolic rate per individual by the average individual size for each treatment. Calculate P:R as GP / R after size normalization. 

```{r}
PRdata$volume.mm3<-size$mean.vol.mm3[match(PRdata$Treatment, size$treatment)]

PRdata<-PRdata%>%
  mutate(R.nmol.mm3.min=R.nmol.org.min/volume.mm3)%>%
  mutate(P.nmol.mm3.min=P.nmol.org.min/volume.mm3)%>%
  mutate(GP.nmol.mm3.min=GP.nmol.org.min/volume.mm3)%>%
  mutate(ratio.size=GP.nmol.mm3.min/(R.nmol.mm3.min*-1))
```

We will need to remove the outlier at 0 and the outliers above 4.
```{r}
PRdata<-PRdata%>%
  mutate(ratio.size=if_else(ratio.size>4, NA, ratio.size))%>%
  mutate(ratio.size=if_else(ratio.size<0.1, NA, ratio.size))
```

### Respiration     

Run ANOVA.  
```{r}
Rmodel2<-PRdata%>%
  filter(!is.na(R.nmol.mm3.min))%>%
  aov(R.nmol.mm3.min~Treatment*Temperature, data=.)

summary(Rmodel2)
```

Significantly lower R when size normalized in larvae from high temperature. 

Calculate % change in R in high temperature larvae. 
```{r}
PRdata%>%
  group_by(Treatment)%>%
  summarise(mean(R.nmol.mm3.min*-1, na.rm=TRUE))%>%
  rename(mean=`mean(R.nmol.mm3.min * -1, na.rm = TRUE)`)%>%
  pivot_wider(names_from=Treatment, values_from=mean)%>%
  mutate(perc_difference=((High-Ambient)/Ambient)*100)


```


Check assumptions of model for residual normality and variance. Passes assumptions.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Rmodel2))
leveneTest(residuals(Rmodel2)~Treatment*Temperature, data=PRdata)
```

### Net Photosynthesis   

Run ANOVA.  
```{r}
Pmodel2<-aov(P.nmol.mm3.min~Treatment*Temperature, data=PRdata)
summary(Pmodel2)
```

Check assumptions of model for residual normality and variance.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(Pmodel2))
leveneTest(residuals(Pmodel2)~Treatment*Temperature, data=PRdata)
```

### Gross Photosynthesis  

Run ANOVA.  
```{r}
GPmodel2<-aov(GP.nmol.mm3.min~Treatment*Temperature, data=PRdata)
summary(GPmodel2)
```

Check assumptions of model for residual normality and variance. Passes.    
```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(GPmodel2))
leveneTest(residuals(GPmodel2)~Treatment*Temperature, data=PRdata)
```

### P:R    

Run ANOVA.  
```{r}
PRmodel2<-PRdata%>%
  filter(!is.na(ratio.size))%>%
  
  aov(ratio.size~Treatment*Temperature, data=.)

summary(PRmodel2)
```

Check assumptions of model for residual normality and variance. Meets assumptions.  

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(PRmodel2))

PRdata%>%
  filter(!is.na(ratio.size))%>%
  leveneTest(residuals(PRmodel2)~Treatment*Temperature, data=.)
```

### Generate summary tables of mean and SE for all variables  

Generate summary of all respiration data.  

```{r}
summary<-PRdata%>%
  group_by(Temperature, Treatment)%>%
  summarise(N=length(R.nmol.org.min),
            Mean_R=mean(R.nmol.org.min*-1, na.rm=TRUE), 
            SE_R=sd(R.nmol.org.min, na.rm=TRUE)/sqrt(length(R.nmol.org.min)),
            Mean_P=mean(P.nmol.org.min, na.rm=TRUE), 
            SE_P=sd(P.nmol.org.min, na.rm=TRUE)/sqrt(length(P.nmol.org.min)),
            Mean_GP=mean(GP.nmol.org.min, na.rm=TRUE), 
            SE_GP=sd(GP.nmol.org.min, na.rm=TRUE)/sqrt(length(GP.nmol.org.min)),
            Mean_PR=mean(ratio, na.rm=TRUE), 
            SE_PR=sd(ratio, na.rm=TRUE)/sqrt(length(ratio)))

summary%>%
  write_csv(., "Mcap2021/Output/Respiration/mean_respiration.csv")
```

# **Plotting**  

## Individual normalized data

### Respiration  

Plot data as an interaction plot   
```{r}
r_plot<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(R.nmol.org.min), sd=sd(R.nmol.org.min), N=length(R.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean*-1, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=(mean*-1)-se, ymax=(mean*-1)+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(0.5, -0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    ylim(0, 0.05)+
    geom_text(x=1.5, y=0.007, label="Treatment p=0.782", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.01, label="Temperature p=0.129", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.004, label="Interaction p=0.614", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot

```

Plot data as a box plot   
```{r}
r_plot_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = (-1*R.nmol.org.min), group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
  
  #geom_label(aes(label=paste(Run, Chamber.ID)), position=position_dodge(0.3))+
  
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot_2
```

### Photosynthesis (Net)  

Plot data as an interaction plot  
```{r}
p_plot<-PRdata %>%
    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min), sd=sd(P.nmol.org.min), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    geom_text(x=1.5, y=0.007, label="Treatment p=0.097", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.01, label="Temperature p=0.339", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.004, label="Interaction p=0.678", color="darkgray", size=6)+
    ylim(0, 0.05) +
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot

```

Plot data as a box plot   
```{r}
p_plot_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = P.nmol.org.min, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot_2
```

### Photosynthesis (Gross)  

Plot data as an interaction plot   
```{r}
gp_plot<-PRdata %>%
    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(GP.nmol.org.min), sd=sd(GP.nmol.org.min), N=length(GP.nmol.org.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    ylim(0, 0.05) +
    geom_text(x=1.5, y=0.007, label="Treatment p=0.348", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.01, label="Temperature p=0.093", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.004, label="Interaction p=0.479", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot

```

Plot data as a box plot   
```{r}

gp_plot_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = GP.nmol.org.min, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot_2
```

### P:R Ratio  

Plot data as an interaction plot     
```{r}
pr_plot<-PRdata %>%
  
    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(ratio, na.rm=TRUE), sd=sd(ratio, na.rm=TRUE), N=length(ratio), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(0.8, 3), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    ylim(1, 3) +
    geom_text(x=1.5, y=1.28, label="Treatment p=0.054", color="darkgray", size=6)+
    geom_text(x=1.5, y=1.4, label="Temperature p=0.484", color="darkgray", size=6)+
    geom_text(x=1.5, y=1.15, label="Interaction p=0.926", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=14)
      ); pr_plot

```

Plot data as a box plot   
```{r}

pr_plot_2<-PRdata %>%

    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = ratio, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot_2

```


### Generate final figure  

```{r}
full_fig<-plot_grid(r_plot, p_plot, gp_plot, pr_plot, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.4), label_y=1, align="h", labels=c("A", "B", "C", "D"), label_size=18)

ggsave(filename="Mcap2021/Figures/Respiration/respirometry_fig.png", plot=full_fig, dpi=500, width=18, height=6, units="in")
```

## Size normalized data

### Respiration  

Plot data as an interaction plot   
```{r}
r_plot_size<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(R.nmol.mm3.min, na.rm=TRUE), sd=sd(R.nmol.mm3.min, na.rm=TRUE), N=length(R.nmol.mm3.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = (-1*mean), group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=(-1*mean)-se, ymax=(-1*mean)+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(0.5, -0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    ylim(0, 0.95)+
    geom_text(x=1.5, y=0.10, label="Treatment p=0.034", color="black", size=6)+
    geom_text(x=1.5, y=0.00, label="Temperature p=0.120", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.05, label="Interaction p=0.519", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot_size

```

Plot data as a box plot   
```{r}
r_plot_size_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = abs(R.nmol.mm3.min), group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    #geom_label(aes(label=paste(Run, Chamber.ID)), position=position_dodge(0.3))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot_size_2
```

### Photosynthesis (Net)  

Plot data as an interaction plot  
```{r}
p_plot_size<-PRdata %>%
    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.mm3.min), sd=sd(P.nmol.mm3.min), N=length(P.nmol.mm3.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    geom_text(x=1.5, y=0.1, label="Treatment p=0.620", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.00, label="Temperature p=0.334", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.05, label="Interaction p=0.625", color="darkgray", size=6)+
    ylim(0, 0.95) +
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot_size

```

Plot data as a box plot   
```{r}
p_plot_size_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = P.nmol.mm3.min, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Net P (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); p_plot_size_2
```

### Photosynthesis (Gross)  

Plot data as an interaction plot   
```{r}
gp_size_plot<-PRdata %>%
    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(GP.nmol.mm3.min), sd=sd(GP.nmol.mm3.min), N=length(GP.nmol.mm3.min), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    ylim(0, 2) +
    geom_text(x=1.5, y=0.24, label="Treatment p=0.133", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.0, label="Temperature p=0.089", color="darkgray", size=6)+
    geom_text(x=1.5, y=0.12, label="Interaction p=0.398", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_size_plot

```

Plot data as a box plot   
```{r}
gp_plot_size_2<-PRdata %>%
    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = GP.nmol.mm3.min, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Gross P (nmol ", O[2], " mm"^-3, "min"^-1, ")")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); gp_plot_size_2
```

### P:R Ratio  

Plot data as an interaction plot     
```{r}
pr_size_plot<-PRdata %>%

    dplyr::group_by(Treatment, Temperature)%>%
    dplyr::summarise(mean=mean(ratio.size, na.rm=TRUE), sd=sd(ratio.size, na.rm=TRUE), N=length(ratio.size), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Temperature), y = mean, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=1, linetype="dashed", color="black", size=0.75)+
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Temperature)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(0.8, 3), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    ylim(1, 2.8) +
    geom_text(x=1.5, y=1.1, label="Treatment p=0.054", color="darkgray", size=6)+
    geom_text(x=1.5, y=1.2, label="Temperature p=0.484", color="darkgray", size=6)+
    geom_text(x=1.5, y=1, label="Interaction p=0.926", color="darkgray", size=6)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=14)
      ); pr_size_plot

```

Plot data as a box plot   
```{r}
pr_plot_size_2<-PRdata %>%

    group_by(Treatment, Temperature)%>%
    ggplot(., aes(x = as.factor(Temperature), y = ratio.size, group=interaction(Treatment, Temperature))) +
    #geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_boxplot(aes(color=Treatment, group=interaction(Treatment, Temperature)), outlier.size = 0, position = position_dodge(0.85), lwd=1) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Temperature)), pch = 21, size=4, position = position_jitterdodge(0.3)) + 
    xlab("Temperature") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("P:R")))) +
    #scale_y_continuous(limits=c(-0.03, 0.005), labels = scales::number_format(accuracy = 0.01, decimal.mark = '.'))+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot_size_2
```

### Generate final figure  

```{r}
full_fig_size<-plot_grid(r_plot_size, p_plot_size, gp_size_plot, pr_size_plot, ncol=4, nrow=1, rel_heights= c(1,1,1,1), rel_widths = c(1,1,1,1.4), label_y=1, align="h", labels=c("A", "B", "C", "D"), label_size=18)

ggsave(filename="Mcap2021/Figures/Respiration/respirometry_size_fig.png", plot=full_fig_size, dpi=500, width=18, height=6, units="in")
```


