---
title: Larval settlement analysis and plotting
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiontic Integration 2021 settlement data. 

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```

# **Plotting** 

Load dataset, collected as number of larvae settled (out of 100) over time (translates to a %).  

```{r}
settle<-read.csv("Mcap2021/Data/Settlement.csv")

settle$Date<-as.factor(settle$Date)
settle$DateTime<-as.POSIXct(settle$DateTime, tz="HST", format="%m/%d/%y")
```

Plot data as mean larval concentration over time.  

```{r}
settle_plot<-settle %>%
    dplyr::group_by(Date, Treatment)%>%
    dplyr::summarise(mean=mean(Settled), sd=sd(Settled), N=length(Settled), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(Date), y = mean, group=interaction(Date, Treatment))) +
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, Date)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, Date)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Date") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("% Settlement")))) +
    #geom_text(x=3, y=16, label="Treatment p=0.758", color="darkgray", size=8)+
    #geom_text(x=3, y=14, label="Time p<0.001", color="black", size=8)+
    ylim(0, 20)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); settle_plot

#ggsave("Mcap2021/Figures/Settlement.png", settle_plot, dpi=300, w=8.5, h=5, units="in")
```

Plot all data with loess line.    

```{r}
settle_plot2<-settle %>%
    
    ggplot(., aes(x = DateTime, y = Settled, group=Treatment)) +
    geom_point(aes(fill=Treatment), pch = 21, size=4, position = position_jitterdodge(dodge.width=0.05), alpha=0.3) + 
    geom_smooth(aes(group=Treatment, color=Treatment), stat="smooth", lwd=2)+
    xlab("Date") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("% Settlement")))) +
    scale_x_datetime(date_breaks="1 day",minor_breaks=waiver(),labels=date_format("%m-%d"))+
    geom_text(x=as.POSIXct("2021-06-24"), y=26, label="Treatment p=0.758", color="darkgray", size=5)+
    geom_text(x=as.POSIXct("2021-06-24"), y=28, label="Time p<0.001", color="black", size=5)+
    ylim(0, 30)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); settle_plot2

ggsave("Mcap2021/Figures/Settlement.png", settle_plot2, dpi=300, w=5, h=5, units="in")
```

# **Analysis**  

Build linear mixed effect model and examine for settlement.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
settlemodel<-lmer(Settled~Treatment*Date + (1|Tank:Treatment:Date), data=settle) #run as random
anova(settlemodel, type="II")
summary(settlemodel)
```

Check assumptions of model for residual normality and variance. Passes assumptions.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(settlemodel))
leveneTest(residuals(settlemodel)~Treatment*Date, data=settle)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(settlemodel, ~Treatment|Date)
cld(emm)
```