---
title: Larval survival analysis and plotting
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for Symbiontic Integration 2021 survival data. 

# **Setup**  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('effects')
library('emmeans')
library('multcomp')
```

# **Plotting**  

Load dataset, collected as larval concentration over time.  

```{r}
survival<-read.csv("Mcap2021/Data/Survivorship.csv")

survival$Date<-as.factor(survival$Date)
```

Plot data as mean larval concentration over time.  

```{r}
surv_plot<-survival %>%
    dplyr::group_by(HoursExposure, Treatment)%>%
    dplyr::summarise(mean=mean(Larvae.per.mL), sd=sd(Larvae.per.mL), N=length(Larvae.per.mL), se=sd/sqrt(N))%>%
    ggplot(., aes(x = as.factor(HoursExposure), y = mean, group=interaction(HoursExposure, Treatment))) +
    geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, HoursExposure)), pch = 21, size=6, position = position_dodge(0.2)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Treatment, HoursExposure)), width=0, linetype="solid", position=position_dodge(0.2), size=1.3, color="black")+
    xlab("Hours Exposure") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Larvae mL" ^-1)))) +
    geom_text(x=3, y=0.7, label="Treatment p=0.806", color="darkgray", size=8)+
    geom_text(x=3, y=0.6, label="Date p<0.001", color="black", size=8)+
    ylim(0.5, 1.5)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); surv_plot

ggsave("Mcap2021/Figures/Survival.png", surv_plot, dpi=300, w=8.5, h=5, units="in")
```

Plot with individual data points and mean line overlayed on top. 
```{r}
surv_plot2<-survival %>%
    
    ggplot(., aes(x = as.factor(HoursExposure), y = Larvae.per.mL, group=interaction(HoursExposure, Treatment))) +
    #geom_line(position=position_dodge(0.2), size=1.3, aes(color=Treatment, group=Treatment)) +
    geom_point(aes(fill=Treatment, group=interaction(Treatment, HoursExposure)), pch = 21, size=4, position = position_jitterdodge(0.2), alpha=0.2) + 
    geom_smooth(aes(group=Treatment, color=Treatment), stat="smooth")+
    xlab("Hours Exposure") + 
    scale_fill_manual(name="Rearing Treatment", values=c("blue","red"))+
    scale_color_manual(name="Rearing Treatment", values=c("blue","red"))+
    ylab(expression(bold(paste("Larvae mL" ^-1)))) +
    geom_text(x=3.5, y=0.0, label="Treatment p=0.806", color="darkgray", size=5)+
    geom_text(x=3.5, y=0.1, label="Date p<0.001", color="black", size=5)+
    ylim(0, 2.2)+
    theme_classic() + 
    theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); surv_plot2

ggsave("Mcap2021/Figures/Survival.png", surv_plot2, dpi=300, w=5, h=5, units="in")
```



# **Analysis**  

Build linear mixed effect model and examine for survival.        
```{r, results=TRUE, warning=FALSE, message=FALSE}
survmodel<-lmer(Larvae.per.mL~Treatment*Timepoint + (1|Tank:Treatment), data=survival) #run as random
anova(survmodel, type="II")
summary(survmodel)
```

Check assumptions of model for residual normality and variance. Passes assumptions.    

```{r, results=TRUE, warning=FALSE, message=FALSE}
qqPlot(residuals(survmodel))
leveneTest(residuals(survmodel)~Treatment*Timepoint, data=survival)
```

Conduct post hoc test.  

```{r}
emm<-emmeans(survmodel, ~Treatment|Timepoint)
cld(emm)
```