---
title: "Temperature analysis"
author: "Ariana S Huffmyer"
date: "6/24/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Set Up  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library("car") #levenes test
library("ggplot2") #plotting
library("plotrix") #plotting
library("reshape2") #data shaping
require("gridExtra") #Arrange Plots for output
require("plyr") #for ddply
require("dplyr")
require("utils")
library("tidyverse")
library("lubridate")
library("scales")
```

Temperature treatments during experiment.  

Plot temperature data from loggers in conicals and view averages of each treatment (Script and data from Hancock et al. 2021).  

```{r}
temps<-read.csv("Mcap2021/Data/Environmental/Temperature/Temperatures.csv")

temps<-temps%>%
  filter(!Temp>35)%>% #remove artifical values
  filter(!Temp<24) #remove artifical values

temps$DateTime<-as.POSIXct(temps$DateTime, format="%m/%d/%y %H:%M")

temp_plot<-ggplot(temps)+
     geom_vline(xintercept=as.numeric(temps$DateTime[1]), linetype="dotted", color="black", size=0.75)+
    geom_vline(xintercept=as.numeric(temps$DateTime[2920]), linetype="dashed", color="black", size=0.75)+
     geom_vline(xintercept=as.numeric(temps$DateTime[5500]), linetype="solid", color="black", size=0.75)+
    geom_point(aes(DateTime,Temp,color=Treatment),alpha=0.3,size=1)+
     scale_color_manual(values=c("blue","red"))+
     #geom_smooth(aes(DateTime,Temp,color=Treatment),se=FALSE,span=0.15,size=1)+
     scale_x_datetime(date_breaks="2 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
     geom_text(x=as.POSIXct("2021-06-14 18:00:00"), y=31.5, label="*", size=10)+
     geom_text(x=as.POSIXct("2021-06-17 18:00:00"), y=31.7, label="+", size=9)+
     ylim(24.5, 32)+
     ylab(expression(bold("Temperature (°C)")))+
     xlab(expression(bold("Date")))+
     theme_classic() + 
     theme(
      legend.position="none",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12), 
      plot.margin = margin(1,1,1,1, unit="cm")
      ); temp_plot

ggsave(filename="Mcap2021/Figures/Temperatures.png", plot=temp_plot, dpi=300, width=9, height=5, units="in")
```

Display mean temperatures during embryo rearing period. 
```{r}
temps%>% filter(DateTime < as.POSIXct("2021-06-15 08:00:00")) %>%dplyr::summarise(mean=mean(Temp),sd=sd(Temp),se=std.error(Temp), n=length(Temp))
```

Display mean temperatures during larval period. 
```{r}
temps%>% filter(DateTime > as.POSIXct("2021-06-15 08:00:00"))%>% filter(DateTime < as.POSIXct("2021-06-18 18:30:00"))%>% group_by(Treatment)%>%dplyr::summarise(mean=mean(Temp),sd=sd(Temp),se=std.error(Temp), n=length(Temp))
```

Display mean temperatures during settlement period. 
```{r}
temps%>% filter(DateTime > as.POSIXct("2021-06-19 08:00:00"))%>% group_by(Treatment)%>%dplyr::summarise(mean=mean(Temp),sd=sd(Temp),se=std.error(Temp), n=length(Temp))
```

Display values during exposure period .
```{r}
temps%>% filter(DateTime > as.POSIXct("2021-06-15 08:00:00"))%>% filter(DateTime < as.POSIXct("2021-06-18 18:30:00"))%>% group_by(Treatment)%>%dplyr::summarise(max=max(Temp), n=length(Temp))
```







